---
title: "Linear Regression Modeling of Police Shootings"
author: "Jordan Tran"
output: pdf_document
---

```{r setup, include=FALSE}
# knit options
knitr::opts_chunk$set(echo = F,
                      results = 'markup',
                      fig.width = 4,
                      fig.height = 3,
                      fig.align = 'center',
                      message = F,
                      warning = F)

# packages
library(MASS)
library(leaps)
library(tidyverse)
library(modelr)
library(tidymodels)
library(faraway)
```
# Dataset
## Sources 

__Demographic data__ is from the 2014 Census

__Expenditure data__ is from California State Controller City Financial Transactions Report

__Police shooting data__ is from the Washington Post's database
```{r , echo = FALSE}
# solution
policeshooting_nodeaths <- read.csv(file = 'data/complete_county_df2.csv')
policeshootings_withdeaths <- read.csv(file = 'data/complete_county_df3.csv')
policeshooting_nodeaths_clean <- subset(policeshooting_nodeaths, select = -c(Count, Police_Funding, Population, County, X,CommunityDev ))
policeshootings_withdeaths_clean <- subset(policeshootings_withdeaths, select = -c(Count, Police_Funding,
                                                                                   Population, County, X,CommunityDev ))[, c(8, 4, 1, 2, 6, 3, 7, 5,9)]
head(policeshootings_withdeaths_clean)
```
## Variables

__ShootingRatio__ - Total Police Shooting Victims divided by County's Population

__Population_per_Square_Mile__ - Population average per square mile of County

__College_Education__ - Percentage of County Population with a Bachelor's Degree or Higher

__Minority__ - Percentage of County Population that is not White

__Below_Poverty__ - Percentage of County Population that is below poverty line

__IncomePerCapita__ - Income per Capita of County Population

__IncomeMedian__ - Median income of County Population

__Police_Funding_Per_Person__ - Total Police Expenditures of County divided by County's Population

__CommunityDev_Per_Person__ - Total Community Development Expenditures of County divided by County's Population

# Model Selection
```{r fig.align="center",fig.width = 10, fig.height= 6}
# Response vs Potential Predictors
policeshooting_nodeaths_clean %>%
  pivot_longer(cols = c(College_Education, Below_Poverty, Minority, Police_Funding_Per_Person, CommunityDev_Per_Person, IncomeMedian, IncomePerCapita, Population_per_Square_Mile)) %>%
  ggplot(aes(x = value, y = ShootingRatio)) +
  facet_wrap(~ name, scales = 'free_x') +
  geom_point() +
  geom_smooth(method = 'loess', formula = 'y ~ x', se = F, span = 1)
```
```{r fig.align="center",fig.width = 5, fig.height= 5}
#See what is skewed
stripchart(data.frame(scale(policeshooting_nodeaths_clean)), method ="jitter", las=2,
vertical=TRUE)
```
```{r , echo = FALSE}
require(leaps)
#Find default search strategy
b <- regsubsets(ShootingRatio~ College_Education + Minority + Police_Funding_Per_Person + Below_Poverty + IncomeMedian + CommunityDev_Per_Person +IncomePerCapita + Population_per_Square_Mile,data=policeshooting_nodeaths_clean)
rs <- summary(b)
rs$which
```
```{r , echo = FALSE}
#Want to minimize AIC
AIC <- 50*log(rs$rss/50) + (2:8)*2
plot(AIC ~ I(1:8), ylab="AIC", xlab="Number of Predictors")
```
```{r , echo = FALSE}
plot(2:9,rs$adjr2,xlab="No. of Parameters",ylab="Adjusted R-square")
```
```{r , echo = FALSE}
#Want small p and CP around or less than p
plot(2:9,rs$cp,xlab="No. of Parameters",ylab="Cp Statistic")
abline(0,1)
```
```{r , echo = FALSE}
# Linear Regression Model
policefit1 <- lm(ShootingRatio ~ College_Education + Minority + Police_Funding_Per_Person + Below_Poverty, data=policeshooting_nodeaths_clean)
summary(policefit1)
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
augment(policefit1, policeshooting_nodeaths_clean) %>%
  pivot_longer(cols = c(.fitted, College_Education, Minority, Police_Funding_Per_Person, Below_Poverty)) %>%
  ggplot(aes(y = .resid, x = value)) +
  facet_wrap(~ name, scales = 'free_x') +
  geom_point() +
  geom_hline(aes(yintercept = 0)) +
  geom_smooth(method = 'loess', formula = 'y ~ x', se = F, span = 1)
```

```{r , echo = FALSE}
#The Box–Cox method is a popular way to determine a transformation on the
#response. It is designed for strictly positive responses and chooses the transformation
#to find the best fit to the data
boxcox(policefit1, plotit=T)
boxcox(policefit1, plotit=T, lambda=seq(-.5,.5,by=0.1))
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
# Response vs Potential Predictors
policeshooting_nodeaths_clean %>%
  pivot_longer(cols = c(College_Education, Below_Poverty, Minority, Police_Funding_Per_Person, CommunityDev_Per_Person, IncomeMedian, IncomePerCapita, Population_per_Square_Mile)) %>%
  ggplot(aes(x = value, y = log(ShootingRatio))) +
  facet_wrap(~ name, scales = 'free_x') +
  geom_point() +
  geom_smooth(method = 'loess', formula = 'y ~ x', se = F, span = 1)
```
```{r , echo = FALSE}
require(leaps)
b <- regsubsets(log(ShootingRatio)~ College_Education + Minority + Police_Funding_Per_Person + Below_Poverty + IncomeMedian + CommunityDev_Per_Person +IncomePerCapita + Population_per_Square_Mile,data=policeshooting_nodeaths_clean)
rs <- summary(b)
rs$which
```
```{r , echo = FALSE}
AIC <- 50*log(rs$rss/50) + (2:8)*2
plot(AIC ~ I(1:8), ylab="AIC", xlab="Number of Predictors")
```
```{r , echo = FALSE}
plot(2:9,rs$adjr2,xlab="No. of Parameters",ylab="Adjusted R-square")
```
```{r , echo = FALSE}
plot(2:9,rs$cp,xlab="No. of Parameters",ylab="Cp Statistic")
abline(0,1)
```
```{r , echo = FALSE}
# Linear Regression Model
policefit2 <- lm(log(ShootingRatio) ~ College_Education + Minority + Police_Funding_Per_Person + Below_Poverty  + CommunityDev_Per_Person, data=policeshooting_nodeaths_clean)
summary(policefit2)
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
#Constant Variance fixed?
augment(policefit2, policeshooting_nodeaths_clean) %>%
  pivot_longer(cols = c(.fitted, College_Education, Minority, Police_Funding_Per_Person, Below_Poverty, CommunityDev_Per_Person)) %>%
  ggplot(aes(y = .resid, x = value)) +
  facet_wrap(~ name, scales = 'free_x') +
  geom_point() +
  geom_hline(aes(yintercept = 0)) +
  geom_smooth(method = 'loess', formula = 'y ~ x', se = F, span = 1)
```
```{r , echo = FALSE}
augment(policefit2, policeshooting_nodeaths_clean) %>%
  ggplot(aes(sample = .resid)) +
  geom_qq() +
  geom_qq_line()
```

```{r fig.align="center",fig.width = 10, fig.height= 6}
#Influential Points
p_caseinf <- augment(policefit2, policeshooting_nodeaths) %>%
  pivot_longer(cols = c(.resid, .hat, .cooksd)) %>%
  ggplot(aes(x = County, y = value)) +
  facet_wrap(~ name, scales = 'free_y', nrow = 3) + # looks better with vertical faceting
  geom_point() +
  geom_hline(aes(yintercept = 0)) + # add line at zero
  theme(axis.text.x = element_text(angle = 90, vjust = 0.25)) + # rotates and aligns labels
  labs(x = '', y = '') 

p_caseinf
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
#Influential Points
h <- lm.influence(policefit2)$hat
names(h) <- policeshooting_nodeaths$County
rev(sort(h))
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
#Dropped SF. Current Model is still best
b<-regsubsets(log(ShootingRatio)~.,data=policeshooting_nodeaths_clean[-c(29),])
rs <- summary(b)
rs$which[which.max(rs$adjr),]
```
```{r , echo = FALSE}
# Linear Regression Model
#Reduced Model is better. Police Funding has no association when other variables are taken in account?
policefit_test <- lm(log(ShootingRatio) ~ College_Education + Minority + Below_Poverty   + CommunityDev_Per_Person, data=policeshooting_nodeaths_clean)
anova(policefit2,policefit_test)
summary(policefit_test)
cor(policeshooting_nodeaths_clean)
#Plot commun and police fund and shooting ratio. scatter x y funding. color shooting ratio
#take out other predictors and look at partial residuals.
#maybe include pop per square mile
#include all predictors. interpret
#Read Chapter 11
# Applied Stats, Biostats, UC Davis, CSU Boulder, Oregon State, Consulting Req/Group, Katie Coburn, CS Programs (Machine learning, deep learning)
```
```{r fig.align="center",fig.width = 10, fig.height= 6}
augment(policefit_test, policeshooting_nodeaths_clean) %>%
  pivot_longer(cols = c(.fitted, College_Education, Minority, Below_Poverty, CommunityDev_Per_Person)) %>%
  ggplot(aes(y = .resid, x = value)) +
  facet_wrap(~ name, scales = 'free_x') +
  geom_point() +
  geom_hline(aes(yintercept = 0)) +
  geom_smooth(method = 'loess', formula = 'y ~ x', se = F, span = 1)
```
```{r , echo = FALSE}
#T distribution?
augment(policefit_test, policeshooting_nodeaths_clean) %>%
  ggplot(aes(sample = .resid)) +
  geom_qq() +
  geom_qq_line()
```
